---
title: "Internel_data_cleansing"
author: "Michiel van Eldik"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

# __0. Prologue__
```{r message=FALSE}
library(dplyr)
library(kableExtra)
library(car)
library(lubridate)
```

```{r Load data}
input <- read.csv("sao_paulo_state_dataset.csv")
```

``` {r create variable}
geo_sao_paolo_df <- input
```

# __1. Text data pre processing__

* First clean
* Then derived variables

## __1.1. Standardization__

* lowercase
* Change encoding to English-letters only (ASCII)

``` {r standardization}
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(
    review_comment_message = tolower(
                                iconv(
                                  review_comment_message, 
                                  to = "ASCII//TRANSLIT"
                                     )
                                    ),
    review_comment_title = tolower(
                              iconv(
                                review_comment_title,
                                to = "ASCII//TRANSLIT"
                                   )
                                  )
        )
```

## 1.2. Punctuation
I do want to keep track of these. It's interesting to know about

* Question marks
* Exclamation marks


``` {r }
library(reticulate)
use_python("/usr/bin/python3.8")
use_virtualenv("/home/mitchy123/Documents/Lemmings/venv")
spacy <- import("spacy")
```

``` {python  }
#import pandas as pd
import spacy
nlp = spacy.load("pt_core_news_sm")
```




``` {r }
# repl_python()
# conda_create("r-reticulate")
# py_install("pandas")
# py_install("spacy")
```

``` {r }
library(reticulate)
use_python("/usr/bin/python3.8")
# virtualenv_create("r-reticulate")
# virtualenv_install("r-reticulate", "scipy")
# virtualenv_install("r-reticulate", "spacy")
use_virtualenv("/home/mitchy123/Documents/Lemmings/venv/bin/python")
scipy <- import("pandas")
spacy <- import("spacy")
```

``` {python  }
#import pandas as pd
import spacy
```

``` {python  }
f = pd.read_csv('sao_paulo_state_dataset.csv', index_col=0)
```

``` {python }
print(f[0:4])
```

``` {python }
nlp = spacy.load("pt_core_news_sm")
```





``` {python }
def foo(x):
  if x == 0:
    print("hoi")
  else:
    print("hi")

foo(20)
``` 



## 1.3. Numbers


## 1.4. Misc weird stuff

## 1.4. Tokenization 

## 1.5. Stop words


Use Portuguese stopwords library

## 1.6. Lemmatization 





## explore text data
Basically I want an answer to whether which comments are noise and which
ones can be taken seriously.

### words with one or multiple expression signs (?, !, *, ...)

``` {r }
# is there a dot?
try <- geo_sao_paolo_df[grep(".", geo_sao_paolo_df$review_comment_message),]

try <- geo_sao_paolo_df[grep("!", geo_sao_paolo_df$review_comment_message),]

```

### Comments with numbers 

### Words with no meaning ()

### single words

### short comments vs actual stories that are useful 
length > xxx

# Dates data 
``` {r }
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(
    review_creation_date = as.Date(review_creation_date,
                                   format = "%Y-%m-%d %H:%M:%S"),
    review_answer_timestamp = as.Date(review_answer_timestamp,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_purchase_timestamp = as.Date(order_purchase_timestamp, 
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_approved_at = as.Date(order_approved_at,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_delivered_carrier_date = as.Date(order_delivered_carrier_date,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_delivered_customer_date = as.Date(order_delivered_customer_date,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_estimated_delivery_date = as.Date(order_estimated_delivery_date,
                                   format = "%Y-%m-%d %H:%M:%S")
         )
```

time relative variables
``` {r }
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(
    diff_est_deliv = order_estimated_delivery_date - order_delivered_customer_date,
    diff_pur_est  = order_estimated_delivery_date - order_purchase_timestamp,
    diff_pur_deliv = order_delivered_customer_date - order_purchase_timestamp,
    diff_rev_crea_ans = review_creation_date - review_answer_timestamp,
    diff_rev_est_ans = order_estimated_delivery_date - review_answer_timestamp,
    diff_rev_deliv_ans = order_delivered_customer_date - review_answer_timestamp
         )
```

weekend dummies 
``` {r }
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(
    # show weekday of a date
    review_sent_dow = wday(review_creation_date, label = TRUE),
    review_answer_dow = wday(review_answer_timestamp, label = TRUE),
    
    # Dummy variable weekend yes/no
    review_sent_wknd = ifelse(
      review_sent_dow == 'zo' | review_sent_dow == 'za', 1, 0),
    review_answer_wknd = ifelse(
      review_answer_dow == 'zo' | review_answer_dow == 'za', 1, 0)
         )
```

Year dummies 
``` {r }
geo_sao_paolo_df<- geo_sao_paolo_df %>%
  mutate(
    y_2016 = ifelse(year(order_purchase_timestamp) == '2016', 1, 0),
    y_2017 = ifelse(year(order_purchase_timestamp) == '2017', 1, 0),
    y_2018 = ifelse(year(order_purchase_timestamp) == '2016', 1, 0)
         )
```

# Duplicates in customer cities



# Past behaviour of customer

* previous recensy 
* frequency
* new_yes/no






