---
title: "Derived variables"
author: "Michiel van Eldik"
date: "`r Sys.Date()`"
output:
   rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# __<a href="index.html">Back to main</a>__

<center> <a href="index.html">Back to main</a> </center>


# __1. Load data and libraries__

```{r message = FALSE}
library(lubridate)
library(dplyr)
library(kableExtra)
```

```{r }
input <- read.csv("sp_2_cleaned_text.csv")
geo_sao_paolo_df <- input
```

# __2. Dates data__
## __2.1. Convert to 'date' type__
``` {r }
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(
    review_creation_date = as.Date(review_creation_date,
                                   format = "%Y-%m-%d %H:%M:%S"),
    review_answer_timestamp = as.Date(review_answer_timestamp,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_purchase_timestamp = as.Date(order_purchase_timestamp, 
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_approved_at = as.Date(order_approved_at,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_delivered_carrier_date = as.Date(order_delivered_carrier_date,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_delivered_customer_date = as.Date(order_delivered_customer_date,
                                   format = "%Y-%m-%d %H:%M:%S"),
    order_estimated_delivery_date = as.Date(order_estimated_delivery_date,
                                   format = "%Y-%m-%d %H:%M:%S")
         )
```

## __2.2. Response time-related variables__

``` {r }
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(
    diff_est_deliv = order_estimated_delivery_date - order_delivered_customer_date,
    diff_pur_est  = order_estimated_delivery_date - order_purchase_timestamp,
    diff_pur_deliv = order_delivered_customer_date - order_purchase_timestamp,
    diff_rev_crea_ans = review_creation_date - review_answer_timestamp,
    diff_rev_est_ans = order_estimated_delivery_date - review_answer_timestamp,
    diff_rev_deliv_ans = order_delivered_customer_date - review_answer_timestamp
         )
```

## __2.3. Date characteristics variables__
Weekend dummies. 
``` {r }
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(
    # show weekday of a date
    review_sent_dow = wday(review_creation_date, label = TRUE),
    review_answer_dow = wday(review_answer_timestamp, label = TRUE),
    
    # Dummy variable weekend yes/no
    review_sent_wknd = ifelse(
      review_sent_dow == 'zo' | review_sent_dow == 'za', 1, 0),
    review_answer_wknd = ifelse(
      review_answer_dow == 'zo' | review_answer_dow == 'za', 1, 0)
         )
```

Year dummies.
``` {r }
geo_sao_paolo_df<- geo_sao_paolo_df %>%
  mutate(
    y_2016 = ifelse(year(order_purchase_timestamp) == '2016', 1, 0),
    y_2017 = ifelse(year(order_purchase_timestamp) == '2017', 1, 0),
    y_2018 = ifelse(year(order_purchase_timestamp) == '2016', 1, 0)
         )
```

# __3. Reviews data__
This is an enormously important part of the research project because in this 
part, the target variable _review creation_ is defined. It begs the question:
What can be considered as the creation of a qualitative review?\n

Technically, each case where`review_comment_message` is not null can 
be considered as a case that has created a qualitative review. That definition
imples a single word like 'bom' would qualify as a qualitative review. But 
does review that is not even a phrase really provide information _beyond_ the 
qualitative review? It's an important question because, ultimately, the added
benefit of qualitative reviews is to disambiguate quantitative review.\n

Firstly, a single word _can_ help to disambiguate a qualitative review. E.g.,
two groups of customers from two different cultures differ in average rating by
1 point on a 5-point Likert scale. Now imagine that all of their quantitative 
reviews are accompanied by the word "boa" (highly hypothetical). In this case, 
the qualitative part of the review reveals to us that perhaps these two groups
have different perceptions of how "boa" relates to a 5-point Likert scale. Hence,
even from one-word comments, information can be derived.\n

Given this notion, even single-word texts are included in the 
definition of _review creation_. Nevertheless, the aspect of message length
and word quantities will still be included in the dataset as it may help in 
explaining outcomes.

## __3.1. Message & Title Dummies__
```{r }
# Create dummy for message vs no message
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(review_comment_message = as.character(review_comment_message)) %>%
  mutate(message_length = nchar(review_comment_message)) %>%
  mutate(message_bool = ifelse(message_length == 0, 0, 1)) %>%
  mutate(message_bool = as.integer(message_bool))

# Create a dummy variable for title message
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(review_comment_title = as.character(review_comment_title)) %>%
  mutate(title_length = nchar(review_comment_title)) %>%
  mutate(title_bool = ifelse(title_length == 0, 0, 1)) %>%
  mutate(title_bool = as.integer(title_bool))

# Create a dummy variable for title OR, AND message
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(title_or_message = ifelse(message_bool == 1 | title_bool == 1, 1, 0)) %>%
  mutate(title_and_message = ifelse(message_bool == 1 & title_bool == 1, 1, 0))
```

## __3.2. Top-2-Box discretization__

```{r }
geo_sao_paolo_df <- geo_sao_paolo_df %>%
  mutate(top2box = ifelse(review_score > 3, 1, 0))
```

# __4. Products data__
## __4.1. Product categories in the data__

There is a whole bunch of product categories. 
```{r }
table(geo_sao_paolo_df$product_category_name)

product_cats <- geo_sao_paolo_df %>%
  group_by(product_category_name) %>%
  summarise(
    freq = n()
  )

```

```{r }
product_cats %>%
  kbl() %>%
  kable_paper(full_width = T) %>%
  scroll_box(height = "500px")

```

## __4.2. SEC classifications__
```{r }
print("Work in progress")

```

## __4.3. Perceived risk classification__

```{r }

print("Work in progress")

```


# __5. Individual level customer data__

The individual customer can carry characteristics. Mainly:

* Whether that person had a drastic experience before 
* How recent this experience was 
* Whether there is a record of previous comments

```{r }
print("Work in Progress")
```



# __6. Write file__

```{r }
write.csv(geo_sao_paolo_df, "sp_3_derived_vars.csv")
```